### 2020-06-02 writeup/intro.tex
intro

Understanding negative transfer requires developing tight generalization bounds for both multi-task learning and single-task learning.
In classical Rademacher or VC based theory of multi-task learning \cite{B00,AZ05,M06}, the generalization bounds are usually presented so that the error reduces as the data sizes of all tasks increase.
For instance, the data sizes of all tasks are often assumed to be equal \cite{B00,LPTV09,LPVT11,MPR16}.
On the other hand, uneven data sizes or dominating tasks have been empirically observed as a cause of negative transfer \cite{YKGLHF20}.
%The benefit of learning multi-task representations has also been studied for certain half-spaces \cite{} and sparse regression \cite{}.
When all tasks are sufficiently similar, it is known that adding more labeled data improves the generalization performance for predicting a particular task \cite{WZR20}.
To motivate our study, we describe simulation results for two linear regression tasks solved using a shared linear layer for both tasks and an output layer for each task.
In Figure \ref{fig_model_shift_phasetrans}, we observe that negative transfer occurs as (a) \textit{task similarity}: tasks become more different; (b) \textit{data size}: source/target data size increases.
%\textbf{Task similarity:}
%\textbf{Data sizes:}
%\textbf{Covariate shift:}
Furthermore, MTL performance is negatively affected when (c) \textit{covariate shift}: the covariance matrices of the two tasks become more different.
The observations highlight the need to develop generalization bounds that scale tightly with properties of the data in order to study MTL performance.

@@

\begin{figure}[!t]
	\begin{subfigure}[b]{0.32\textwidth}
		\centering
		\includegraphics[width=0.98\textwidth]{figures/model_shift_phase_transition.pdf}
		\vspace{-0.075in}
		\caption{Task similarity}
		\label{fig_model_shift}
	\end{subfigure}\hfill
	\begin{subfigure}[b]{0.32\textwidth}
		\centering
		\includegraphics[width=0.98\textwidth]{figures/datapoints_phase_transition.pdf}
		\vspace{-0.075in}
		\caption{Data sizes}
		\label{fig_size}
	\end{subfigure}\hfill
	\begin{subfigure}[b]{0.32\textwidth}
		\centering
		\includegraphics[width=0.98\textwidth]{figures/complementary.pdf}
		\vspace{-0.075in}
		\caption{Covariate shift}
		\label{fig_covariate}
	\end{subfigure}
	\caption{%Three takeaways of our theory in Section \ref{sec_insight}.
	We observe that MTL performance
	(a) transitions from positive to negative transfer as \textit{task model distance} increases (Prop. \ref{prop_dist_transition});
	(b) transitions from positive to negative transfer as source/target \text{data size} increases (Prop. \ref{prop_data_size});
	(c) gets worse as \textit{covariate shift} gets more severe (increasing $\lambda$) and source/target \textit{data size} increases (Prop. \ref{prop_covariate}).
	$y$-axis denotes the prediction loss of STL minus MTL.}
	\label{fig_model_shift_phasetrans}
	\vspace{-0.1in}
\end{figure}

>>>>

Understanding negative transfer requires developing tight generalization bounds for both multi-task learning and single-task learning.
In classical Rademacher or VC based theory of multi-task learning \cite{B00,AZ05,M06}, the generalization bounds are usually presented so that the error reduces as the data sizes of all tasks increase.
For example, the sample sizes of all tasks are often assumed to be equal \cite{B00,LPTV09,LPVT11,MPR16}.
On the other hand, uneven sample sizes (or dominating tasks) have been empirically observed to cause of negative transfer \cite{YKGLHF20}.
%The benefit of learning multi-task representations has also been studied for certain half-spaces \cite{} and sparse regression \cite{}.
When all tasks are sufficiently similar, adding more labeled data improves the generalization performance for predicting a particular task \cite{WZR20}.
To motivate our study, Figure \ref{fig_model_shift_phasetrans} shows three examples for two linear regression tasks solved using a shared linear layer for both tasks and an output layer for each task.
We observe that negative transfer occurs as (a) \textit{task similarity}: tasks become more different; (b) \textit{data size}: source/target data size increases.
%\textbf{Task similarity:}
%\textbf{Data sizes:}
%\textbf{Covariate shift:}
Furthermore, MTL performance is negatively affected when (c) \textit{covariate shift}: the covariance matrices of the two tasks become more different.
The observations highlight the need to develop generalization bounds that scale tightly with properties of multiple task data.

@@

\begin{figure}[!t]
	\begin{subfigure}[b]{0.32\textwidth}
		\centering
		\includegraphics[width=0.98\textwidth]{figures/model_shift_phase_transition.pdf}
		\vspace{-0.075in}
		\caption{Task similarity}
		\label{fig_model_shift}
	\end{subfigure}\hfill
	\begin{subfigure}[b]{0.32\textwidth}
		\centering
		\includegraphics[width=0.98\textwidth]{figures/datapoints_phase_transition.pdf}
		\vspace{-0.075in}
		\caption{Sample size}
		\label{fig_size}
	\end{subfigure}\hfill
	\begin{subfigure}[b]{0.32\textwidth}
		\centering
		\includegraphics[width=0.98\textwidth]{figures/complementary.pdf}
		\vspace{-0.075in}
		\caption{Covariate shift}
		\label{fig_covariate}
	\end{subfigure}
	\caption{%Three takeaways of our theory in Section \ref{sec_insight}.
	We observe that MTL performance
	(a) transitions from positive to negative transfer as \textit{task model distance} increases (Prop. \ref{prop_dist_transition});
	(b) transitions from positive to negative transfer as source/target \text{sample size} increases (Prop. \ref{prop_data_size});
	(c) gets worse as \textit{covariate shift} gets more severe (increasing $\lambda$) and source/target \textit{sample size} increases (Prop. \ref{prop_covariate}).
	The $y$-axis measures the loss of STL minus MTL.}
	\label{fig_model_shift_phasetrans}
	\vspace{-0.1in}
\end{figure}
